---
title: "Homework 2 SDS315"
author: "Niva Shirsekar, EID: ns38287"
date: "https://github.com/NivaShrskr/SDS-315"
output: pdf_document
---

```{r global_options, echo=FALSE}
#global options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, tidy = TRUE, fig.width=8, fig.height=6)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#loading in necessary libraries 
library(tidyverse)
library(dplyr)
library(forcats)
library(lubridate)
library(purrr)
library(stringr)
library(readr)
library(tibble)
library(tidyr)
library(kableExtra)
library(rvest)
library(sas7bdat)


#creating function to make tables easier
my_nice_table = function(your_table, vector_nice_names, nice_title) {
  nice = kable(your_table, col.names = vector_nice_names, caption = nice_title)
  return(nice)
}
```

# **Question 1**
```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
profs = read.csv("profs.csv")
#defining ggplot
profs_ggplot <- ggplot(profs)
```

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Create a histogram to display the overall data distribution of course evaluation scores.
profs_ggplot + geom_histogram(aes(x=eval), color = "black", fill = "darkred", binwidth = 0.1) + labs(title = "Distribution of Course Evaluation Scores", x = "Scores on a Scale of 1-5", y = "Frequency") +
        theme_minimal()
```
The distribution of course evaluations is left-skewed, with most scores clustering between 3.5 and 4.5. On the graph, the x-axis represents course evaluation scores, while the y-axis indicates the frequency of professors receiving scores within each range.

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Use side-by-side boxplots to show the distribution of course evaluation scores by whether or not the professor is a native English speaker.
profs_ggplot +   geom_boxplot(aes(x = native, y = eval, fill = native))  +
  scale_fill_manual(values = c("lavender", "pink")) +
  labs(
    title = "Course Evaluation Scores by Native English Speaker",
    x = "Native English Speaker",
    y = "Evaluation Score"
  ) +
  theme_minimal()
```
The box plots reveal that professors who are native English speakers tend to receive higher course evaluation scores compared to non-native English-speaking professors. In the graph, the x-axis indicates whether the professor is a native English speaker, while the y-axis represents the course evaluation scores.

### **Part C**

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Part C. Use a faceted histogram with two rows to compare the distribution of course evaluation scores for male and female instructors.
profs_ggplot +   geom_histogram(aes(x = eval, fill = gender), color = "grey3", binwidth=0.1)  +
  scale_fill_manual(values = c("salmon", "chartreuse4")) +
  labs(
    title = "Comparing distribution of course evaluations by gender",
    x = "Evaluation score",
    y = "Frequency"
  ) + facet_wrap(~ gender, scales = "free_y") +
  theme_minimal()
```
The histograms show that male professors generally received higher course evaluation scores compared to female professors. The x-axis represents the course evaluation scores, while the y-axis indicates the frequency of professors receiving scores within each range.

### **Part D**

```{r echo=FALSE, warning=FALSE, message=FALSE}
profs_ggplot + geom_point(aes(x=beauty,y=eval), color = "tomato") + labs(
    title = "Comparing a Professor's Physical Attractiveness and Their Course Evaluations",
    x = "Attractiveness",
    y = "Evaluation"
  ) 
```
The scatterplot above reveals a very weak positive correlation between professors' physical attractiveness and their course evaluation scores. The x-axis represents the physical attractiveness scale, while the y-axis shows the corresponding course evaluation scores.

# **Question 2**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
bikes = read.csv("bikeshare.csv")
#defining ggplot
bikes_ggplot <- ggplot(bikes)
```

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Create a line graph showing average hourly bike rentals (total) across all hours of the day (hr).
#Wrangling Data
bikes_ggplot <- bikes %>%
  #grouping data by hour
  group_by(hr) %>%
  #summarizing the mean number of bike rentals for each group of hours, ignoring NA values
  summarize(avg_total = mean(total, na.rm = TRUE))
#creating line plot
ggplot(bikes_ggplot, aes(x = hr, y = avg_total)) + geom_line(color = "magenta", size = 1) + labs(title = "Average Hourly Bike Rentals", x = "Hour of the Day", y = "Average Number of Rentals") + theme_minimal() 
```
This line graph depicts the relationship between the hour of the day (x-axis) and the average number of bike rentals (y-axis), revealing a clear bimodal pattern. Peaks occur at 8:00 a.m. and 4:00 p.m., with average rental frequencies of approximately 350 and 460 bikes, respectively. These peaks suggest that most bike rentals coincide with typical commuting times, indicating that people are more likely to bike in the morning or late afternoon. In contrast, the lowest point on the graph is at 4:00 a.m., where bike rentals drop to an average of about 10, reflecting minimal activity during early morning hours. This trend highlights the impact of daily routines and commuting habits on bike rental patterns. 

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Plot B: a faceted line graph showing average bike rentals by hour of the day, faceted according to whether it is a working day (workingday)
bikes_ggplot <- bikes %>%
  #grouping data by hour and workingday
  group_by(hr, workingday) %>%
  #summarizing the mean number of bike rentals for each group of hours, ignoring NA values
  summarize(avg_total = mean(total, na.rm = TRUE))
ggplot(bikes_ggplot, aes(x = hr, y = avg_total, color = factor(workingday))) + 
  geom_line(size = 1) +
  labs(title = "Average Bike Rentals by Hour of the Day", x = "Hour of the Day", y = "Average Number of   Rentals", color = "Working Day") + scale_color_manual(values = c("0" = "orange", "1" = "tan"), labels = c("0" = "No", "1" = "Yes")) +  # Customize labels
  facet_wrap(~ workingday, scales = "free_y", labeller = labeller(workingday = c("0" = "Not Workingday", "1" = "Workingday"))) +  # Customize facet labels
  theme_minimal()
```
The graphs illustrate the average number of bike rentals per hour, differentiated between weekdays and weekends, offering insight into how rental patterns vary based on the day. On weekends, the graph displays a unimodal distribution, with a single peak occurring around noon, where the average number of bike rentals reaches approximately 370. This pattern suggests that weekend bike usage is likely tied to leisure activities, with a preference for mid-day outings. In contrast, the weekday graph exhibits a distinct bimodal distribution, with two prominent peaks. The first peak occurs at 8:00 a.m., with an average of 480 rentals, while the second appears at 4:00 p.m., with a slightly higher average of 520 rentals. These peaks align with typical commuting hours, indicating that weekday bike rentals are driven primarily by work and school schedules.

### **Part C**

```{r echo=FALSE, warning=FALSE, message=FALSE}
bikes_9am <- bikes %>%
  filter(hr==9 ) %>%
  group_by(weathersit, workingday) %>%
  summarize(avg_total = mean(total, na.rm = TRUE))

ggplot(bikes_9am, aes(x = factor(weathersit), y = avg_total, fill = factor(weathersit))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Ridership During 9 AM by Weather Situation", x = "Weather Situation", y = "Avg Ridership", fill = "Weather Situation") + scale_fill_manual(values = c("1" = "plum1", "2" = "plum3", "3" = "plum4"),labels = c("1" = "Clear, Few clouds, Partly cloudy, Partly cloudy", "2" = "Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist","3" = "Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds")) + facet_wrap(~ workingday, scales = "free_y", labeller = labeller(workingday = c("0" = "Not Workingday", "1" = "Workingday"))) + theme_minimal() + theme(legend.position = "bottom")
```
These bar plots illustrate the average bike ridership at 9 a.m. under various weather conditions, comparing weekday and weekend trends. On weekdays, the data reveals that most individuals, approximately 250 riders, are willing to bike as long as the weather remains fair. However, when conditions deteriorate to include light rain or snow, the number of riders drops significantly, with average ridership decreasing to around 140. This sharp decline suggests that weekday bike use is largely utilitarian, such as commuting to work, but is highly sensitive to adverse weather. In contrast, weekend ridership demonstrates a different pattern. People are more inclined to ride their bikes when the weather is sunny or partly cloudy, likely for leisure or recreational purposes. As weather conditions worsen, there is a steady decline in the number of riders, reflecting a greater tendency to avoid biking in inclement weather during non-essential outings.

# **Question 3**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
metro = read.csv("capmetro_UT.csv")

# Recode the categorical variables in sensible, rather than alphabetical, order
metro = mutate(metro,
day_of_week = factor(day_of_week,
levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
month = factor(month,
levels=c("Sep", "Oct","Nov")))

#defining ggplot
metro_ggplot <- ggplot(metro)
```

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
avg_boardings <- metro %>%
  group_by(hour_of_day,day_of_week, month) %>%
  summarize(avg_boardings = mean(boarding, na.rm = TRUE))

ggplot(avg_boardings, aes(x = hour_of_day, y = avg_boardings, color = factor(month))) +
  geom_line(size = .5) + facet_wrap(~ day_of_week, nrow = 2) + scale_color_manual(
    values = c("red", "forestgreen", "cyan"),  
    labels = c("September", "October", "November")  
  ) + 
  labs(
    title = "Average Boardings by Hour of Day, Day of Week, and Month",
    x = "Hour of the Day",
    y = "Average Boardings",
    color = "Month"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")  
  )

```
This line graph illustrates the average boardings by hour of the day for each day of the week, with separate lines for September, October, and November. The hour of peak boardings is broadly consistent across all days, occurring in the morning (around 8 AM) and evening (around 5 PM), reflecting standard commute times for work or school. However, some variations are evident when comparing specific days and months. For instance, Mondays in September have lower average boardings than other weekdays, likely due to reduced activity at the start of the academic year as schools and universities transition into their regular schedules or lingering summer vacations. Similarly, Wednesdays, Thursdays, and Fridays in November show a noticeable decline in boardings, which is likely tied to the Thanksgiving holiday and its impact on schedules, with many people traveling or taking extended breaks. These trends highlight the consistent influence of daily routines on ridership patterns, while also reflecting the effects of seasonal and cultural events.

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(metro) + geom_point(aes(x=temperature,y=boarding,color=weekend), size =.5, alpha = .5) + facet_wrap(~hour_of_day) + theme_minimal() + labs(title="Boardings vs Temperature by Hour of Day and Day Status",x="Temperature (F)",y="Boardings")
```
This scatter plot reveals the relationship between boardings and temperature, with the data faceted by hour of the day and color-coded based on whether the day is a weekend or a weekday. When we hold both the hour of the day and weekend status constant, the plot clearly indicates that temperature has a noticeable effect on ridership. As the temperature increases, boardings generally tend to rise as well, especially during midday hours. This pattern suggests that UT students are more likely to ride the bus in warmer temperatures, likely because the weather is more conducive to outdoor activities or travel. On weekends, the correlation between temperature and ridership is still visible, but it appears less pronounced compared to weekdays. This difference could be due to altered routines on weekends, where students may have different travel needs or flexibility in their schedules. Overall, temperature plays a significant role in determining bus ridership, with warmer weather encouraging higher usage, particularly in the afternoon when temperatures tend to peak.

\newpage


# **Question 4**

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
billboard = read.csv("billboard.csv")
top_10 <- billboard %>%
  mutate(week=1) %>%
  group_by(song, performer) %>%
  summarize(total_weeks = sum(week, na.rm = TRUE)) %>%
  arrange(desc(total_weeks)) %>%
  head(n = 10)

my_nice_table(your_table = top_10, vector_nice_names = c("Song", "Performer", "Total Weeks"), nice_title = "Top 10 Songs by Total Weeks on Billboard Top 100")

```

This table lists the top 10 most popular songs since 1958, ranked by the total number of weeks on the Billboard Top 100 chart. It includes the performer, song title, and total weeks on the chart.

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
mus_div <- billboard %>%
  filter(year != 1958 & year != 2021) %>%
  group_by(year) %>%
  summarize(unique_song = n_distinct(song), .groups = "drop")

ggplot(mus_div, aes(x = year, y = unique_song)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Musical Diversity by Year", x = "Year", y = "Musical Diversity")


```
This line graph illustrates that musical diversity peaked in the late 1960s, followed by a gradual decline until the early 2000s. However, since then, there has been a sharp rise in diversity, with numbers nearly reaching the 1960s peak of 800 by 2020.

### **Part C** 

```{r echo=FALSE, warning=FALSE, message=FALSE}
ten_week_hit <- mutate(billboard, week=1)
ten_week_hit <- group_by(ten_week_hit, song, performer)
ten_week_hit <- summarise(ten_week_hit, total_weeks= sum(week, na.rm = TRUE))
ten_week_hit <- filter(ten_week_hit, total_weeks >= 10)

ten_week_hit <- group_by(ten_week_hit,performer)
ten_week_hit <- summarise(ten_week_hit,songs=length(song))
ten_week_hit <- filter(ten_week_hit,songs >= 30)

ggplot(ten_week_hit) + geom_col(aes(fct_reorder(performer,songs),songs),fill="orange",color="black") + coord_flip() + theme_minimal() + labs(title="Artists with 30 Plus Ten Week Hits",y="Songs",x="Artists")

```

This bar graph displays the 19 artists who have had at least 30 songs chart on the Billboard Top 100 for a minimum of 10 weeks. The x-axis represents the number of ten-week hits, while the y-axis lists the artists.


