---
title: "Homework 3 SDS315"
author: "Niva Shirsekar, EID: ns38287"
date: "https://github.com/NivaShrskr/SDS-315"
output:
  pdf_document:
    keep_tex: true
---

```{r global_options, echo=FALSE}
#global options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, tidy = TRUE, fig.width=8, fig.height=6)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#loading in necessary libraries 
library(tidyverse)
library(dplyr)
library(forcats)
library(lubridate)
library(purrr)
library(stringr)
library(readr)
library(tibble)
library(tidyr)
library(kableExtra)
library(rvest)
library(sas7bdat)
library(mosaic)


#creating function to make tables easier
my_nice_table = function(your_table, vector_nice_names, nice_title) {
  nice = kable(your_table, col.names = vector_nice_names, caption = nice_title)
  return(nice)
}
```

\newpage

# **Question 1**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
gas = read.csv("gasprices.csv")
gas_ggplot = ggplot(gas)
```

### **Part A**

##### Claim: 
Gas stations charge more if they lack direct competition in sight.

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(gas, aes(x = Price)) +
  geom_histogram(fill = "pink", color = "grey6") +
  facet_wrap(~ Competitors) +
  theme_minimal() +
  labs(x = "Price ($)", y = "Frequency") 


boot_comp = do(10000)*diffmean(Price ~ Competitors, data=mosaic::resample(gas))
confint(boot_comp)

```
##### Evidence: 
The histogram shows a similar spread of prices for both gas stations with or without competition in sight. The difference in price between gas stations with and without competition in sight is somewhere between -0.0555 and 0.0076, with 95% confidence. 

##### Reasoning: 
Since this confidence interval includes the value 0, we cannot say that there is a significant difference in price based on if a gas station has competition nearby or not. 

\newpage


### **Part B**

##### Claim: 
The richer the area, the higher the gas prices.

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Calculate the median income
median_income <- median(gas$Income, na.rm = TRUE)

# Create the High_Income variable based on the median
gas <- gas %>%
  mutate(High_Income = ifelse(Income > median_income, "Yes", "No"))

ggplot(gas, aes(x = Price)) +
  geom_histogram(fill = "pink", color = "grey6") +
  facet_wrap(~ High_Income) +
  theme_minimal() +
  labs(x = "Price ($)", y = "Frequency") 

boot_inc <- do(10000) * diffmean(Price ~ High_Income, data = resample(gas))
# Confidence Interval
confint(boot_inc)
```

##### Evidence: 

The histograms show that there seems to be a higher amount of lower-priced gas stations in lower income areas. We considered that the the average incomes that were higher than the median income of about 52k to be high income, and those lower than that value to be considered lower income. The difference in price between gas stations in lower and higher income areas is somewhere between -0.0116 and 0.0527, with 95% confidence. 

##### Reasoning: 
Since this confidence interval includes the value 0, we cannot say that there is a significant difference in price based on if a gas station is in a high income area or not. 

\newpage

### **Part C**

##### Claim: 
Gas stations at stoplights charge more.

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(gas, aes(x = Stoplight, y = Price)) +
  geom_boxplot() +
  labs(title = "Gas Prices by if at a Stoplight or Not", x = "Stoplight", y = "Price ($)") +
  theme_minimal()

boot_stop <- do(10000) * diffmean(Price ~ Stoplight, data = resample(gas))
# Confidence Interval
confint(boot_stop)
```

##### Evidence:  
The box plots show that there seems to be a higher median price in gas stations by a stoplight, at about 1.87, versus the median price of gas stations not near a stoplight at about 1.83. The difference in price between gas stations in near a stoplight and not near a stoplight is somewhere between -0.0382 and 0.0300, with 95% confidence.   

##### Reasoning: 
Since this confidence interval includes the value 0, we cannot say that there is a significant difference in price based on if a gas station is near a stoplight or not. 

\newpage

### **Part D**

##### Claim:  
Gas stations with direct highway access charge more

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(gas, aes(x = Highway, y = Price)) +
  geom_boxplot() +
  labs(title = "Gas Prices by if Direct Access to Highway or Not", x = "Access to Highway", y = "Price ($)") +
  theme_minimal()

boot_high <- do(10000) * diffmean(Price ~ Highway, data = resample(gas))
# Confidence Interval
confint(boot_high)
```

##### Evidence:  
The box plots show that there seems to be a higher median price in gas stations by the highway, at about 1.89, versus the median price of gas stations not near a highway at about 1.84. The difference in price between gas stations in near a stoplight and not near a stoplight is somewhere between 0.0087 and 0.0804, with 95% confidence. 

##### Reasoning: 
We can say with 95% confidence that there is a small difference of 0.0358 between the price of gas stations in lower and high income areas. 

\newpage


### **Part E**

##### Claim: 
Shell charges more than all other non-Shell brands.

```{r echo=FALSE, warning=FALSE, message=FALSE}
shell_gas <- gas %>%
  mutate(Station_Type = ifelse(Brand == "Shell", "Shell", "Non-Shell"))

ggplot(shell_gas, aes(x = Station_Type, y = Price)) +
  stat_summary(fun = "mean", geom = "bar", fill = c("tomato", "lightblue")) +
  labs(
    title = "Average Gas Price: Shell vs Non-Shell",
    x = "Gas Station Type",
    y = "Average Price ($)"
  ) +
  theme_minimal()

boot_type <- do(10000) * diffmean(Price ~ Station_Type, data = resample(shell_gas))
# Confidence Interval
confint(boot_type)
```

##### Evidence:  
The histograms show that there seems to be about the same price between shell and non shell gas stations. The difference in price between gas stations owned by shell and not owned by shell is somewhere between -0.0099 and 0.0647, with 95% confidence. 

##### Reasoning: 

Since this confidence interval includes the value 0, we cannot say that there is a significant difference in price based on if a gas station is owned by shell or not. 

\newpage

# **Question 2**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
class = read.csv("sclass.csv")
#defining ggplot
class_ggplot <- ggplot(class)
```

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
class_filtered <- class %>%
  filter(year == 2011, trim == "63 AMG") %>% 
  mutate(Car_Status = ifelse(condition == "Used", "Used", "Unused"))  

boot_mileage <- do(10000) * diffmean(mileage ~ Car_Status, data = resample(class_filtered))
conf_interval <- confint(boot_mileage)

# Extract the numeric values (lower and upper bounds)
numeric_conf_interval <- conf_interval[, 2:3]

# Round to 4 decimal places
rounded_conf_interval <- round(numeric_conf_interval, 2)
print(rounded_conf_interval)
```

The difference in price between gas stations owned by shell and not owned by shell is somewhere between -1115.47 and	9630.82, with 95% confidence. 

\newpage

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
class_filtered2 <- class %>%
  filter(year == 2014, trim == "550") %>%
  mutate(isBlack = ifelse(color == "Black", TRUE, FALSE))

boot_class <- do(10000) * prop(isBlack ~ 1, data = resample(class_filtered2))
conf_interval_class <- confint(boot_class)

# Extract the numeric values (lower and upper bounds)
numeric_conf_interval_class <- conf_interval_class[, 2:3]

# Round to 4 decimal places
rounded_conf_interval_class <- round(numeric_conf_interval_class, 4)
print(rounded_conf_interval_class)
```

The difference in price between gas stations owned by shell and not owned by shell is somewhere between 0.4168 and	0.4528, with 95% confidence. 

\newpage

# **Question 3**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
nbc = read.csv("nbc_pilotsurvey.csv")
#defining ggplot
nbc_ggplot <- ggplot(nbc)
```

### **Part A**

```{r echo=FALSE, warning=FALSE, message=FALSE}
nbc_shows <- nbc %>%
  filter(Show == "Living with Ed" | Show == "My Name is Earl")

boot_nbc_show <- do(10000) * diffmean(Q1_Happy ~ Show, data = resample(nbc_shows))
confint(boot_nbc_show)
```

#### Question:  
Is there evidence that either the show “Living with Ed” or the show “My Name is Earl” consistently produces a higher mean happiness response among viewers?  

#### Approach: 
I first filtered the original nbc dataset to include only the two shows in question. I then ran a confidence interval with 10000 simulations to determine if one show consistently produces a higher mean happiness response among viewers.   

#### Results:  
I found that the difference in mean happiness between the shows “Living with Ed” and “My Name is Earl” is somewhere between -0.4007 and 0.1036, with 95% confidence.

#### Conclusion:   
Since this confidence interval includes the value 0, we cannot say that there is a significant difference in viewer happiness level from watching one show or the other. This implies that viewers found both series equally as funny.   

\newpage

### **Part B**

```{r echo=FALSE, warning=FALSE, message=FALSE}
nbc_shows2 <- nbc %>%
  filter(Show == "The Biggest Loser" | Show == "The Apprentice: Los Angeles")

boot_nbc_show2 <- do(10000) * diffmean(Q1_Annoyed ~ Show, data = resample(nbc_shows2))
confint(boot_nbc_show2)
```

#### Question:  
Which reality TV show, “The Biggest Loser” or “The Apprentice: Los Angeles", made people feel the most annoyed? 

#### Approach:  
I first filtered the original nbc dataset to include only the two shows in question. I then ran a confidence interval with 10000 simulations to determine if one show consistently produces a higher mean annoyed response among viewers.     

#### Results:   
I found that the difference in mean annoyance between the shows “The Biggest Loser” and “The Apprentice: Los Angeles" is somewhere between -0.5270 and -0.0171, with 95% confidence.  

#### Conclusion:   
We can say with 95% confidence that there is a small difference of 0.0343 between the annoyance levels in the shows “The Biggest Loser” and “The Apprentice: Los Angeles". This means that there was a small, but significant difference in the annoyance levels for the two shows.  

\newpage

### **Part C**


```{r echo=FALSE, warning=FALSE, message=FALSE}
nbc_shows3 <- nbc %>%
  filter(Show == "Dancing with the Stars") %>%
  mutate(new_Q2 = ifelse(Q2_Confusing >= 4, 1, 0))

boot_nbc_show3 <- do(10000) * prop(new_Q2 ~ 1, data = resample(nbc_shows3))
confint(boot_nbc_show3)
```
#### Question:  
What proportion of American TV watchers would we expect to give a response of 4 or greater to the “Q2_Confusing” question in the show Dancing with the Stars?

#### Approach:  
I first filtered the original nbc dataset to include only the the show "Dancing with the Stars". I mutated the dataset to only include the values 4 and five, because any response of 4 or 5 indicated that the survey participant either Agreed (4) or Strongly Agreed (5) that “Dancing with the Stars” was a confusing show. I then ran a confidence interval with 10000 simulations to determine proportion of American TV watchers would give a response of 4 or greater to the “Q2_Confusing” question. 

#### Results:   
I found that the difference in proportion in the show "Dancing with the Stars" is somewhere between -0.03867 and 0.1160, with 95% confidence.  

#### Conclusion:   
Since this confidence interval includes the value 0, we cannot say that there is a significant proportion of viewers who were confused from watching "Dancing with the Stars".  This indicates that most people are not confused by the show


\newpage


# **Question 4**

```{r echo=FALSE, warning=FALSE, message=FALSE}
#reading in dataset
ebay = read.csv("ebay.csv")
#defining ggplot
ebay_ggplot <- ggplot(ebay)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
ebay <- ebay %>%
  mutate(rev_ratio = rev_after / rev_before)

boot_ebay <- do(10000) * diffmean(rev_ratio ~ adwords_pause, data = resample(ebay))
confint(boot_ebay)
```

#### Question:
We are trying to determine whether EBay's paid advertising on Google’s search platform increases its revenue compared to a control group where ads were paused.

#### Approach:
To answer this, I calculated the revenue ratio (after-to-before) for each DMA, computed the difference in means between the treatment and control groups, and used a Monte Carlo simulation with 10,000 resamples to generate a 95% confidence interval for the difference.

####  Results:
The observed difference in mean revenue ratios was -0.052, and the 95% confidence interval ranged from -0.091 to -0.013, indicating a small statistically significant difference.

#### Conclusion:
The analysis suggests that EBay’s paid advertising on Google increased its revenue by a small amount, as the treatment group had a small significantly lower revenue ratio, implying the ads had a positive effect.

